Cooperative Control for Narrow Passage Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station

### 2012-10-11 manuscript.tex
Adding more path figures (not using them all, just so everyone can see) and added text

% Packages
\usepackage{color}
\usepackage[pdftex]{hyperref}

% PDF bookmarks
\definecolor{linkCol}{gray}{0.25}
\hypersetup{
    pdftitle={Cooperative Control for Narrow Passage Traversal with an 
Ornithopter Micro Air Vehicle and Lightweight Ground Station},
    pdfauthor={R.C. Julian, C.J. Rose, H. Hu, and R.S. Fearing},
    pdfsubject={},
    pdfkeywords={AAMAS proceedings, multiagent systems, cooperative control, 
particle filter, micro air vehicle, ground station, visual servoing, pose 
estimation, flapping wing, ornithopter, biomimetic},
    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
}

@@

% In the original styles from ACM, you would have needed to
% add meta-info here. This is not necessary for AAMAS 2013  as
% the complete copyright information is generated by the cls-files.
\title{Cooperative Control for Narrow Passage Traversal with an 
Ornithopter Micro Air Vehicle and Lightweight Ground Station}

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
%TODO(humphrey)
The flying platform used in this experiment is the H$^2$Bird flapping wing 
MAV. Built around the Silverlit I-Bird RC flyer power train and clap-fling 
wings, the H$^2$Bird has a wingspan of [40?] cm and has a flight weight of 
13 grams. A tail propeller and servo-controlled tail elevator provide 
control in the yaw and pitch axes. An on-board electronics package comprises 
a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
can operate for approximately 10 minutes.

%TODO footnote for Silverlit I-Bird
%TODO measure wingspan
%TODO ground speed

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
%TODO ryan

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

@@

%----------------------------------------------------------------------------%
\subsection{Ornithoper Flight Control}

\begin{figure}[tb]
\centering
\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
\caption{H$^2$Bird Ornithopter with flight control axes}
\label{fig:h2Bird_axes}
\end{figure}

@@

Due to the high pitch angles experienced by the H$^2$Bird, we use a 
quaternion representation of vehicle pose to avoid the numerical 
instabilities of Euler angle representations. This choice of pose 
representation also lends itself well to remote guidance, as relative body 
and world coordinate angular displacements can be represented compactly 
with quaternion multiplication.~\cite{bowman:reasoning}

Given a reference pose $q$, we define the error $q_e$ as the rotation 
required to reach the reference pose from the measured pose. We then 
linearise the error quaternion to produce body axis angular errors as as 
inputs to the attitude controllers.

@@

\begin{equation}
\label{quat_error}
q_e = q'\otimes q_r
\end{equation}
\begin{equation}
\label{quat_angle}
\alpha = 2acos(q_{e,w})
\end{equation}
\begin{equation}
\label{quat_linearize}
Q_e = \alpha*q_e/sin(\alpha /2)
\end{equation}

%----------------------------------------------------------------------------%
\subsection{Pose Estimation}
%TODO ryan

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
%TODO ryan

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%TODO ryan

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental environment.}
\label{fig:experiment_cartoon}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
%TODO cameron

%----------------------------------------------------------------------------%
\subsection{Ornithopter Flight Control}
\label{sec:flight_control}
%TODO cameron
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
\caption{H$^2$Bird step response}
\label{fig:step_response}
\end{figure}

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. The results of this experiment are depicted in
Figure~\ref{fig:step_response}. The data for the experiment was gathered from
the integration of the H$^2$Bird's built in gyroscope data. We used Matlab to
fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
turning behavior to the step response with the physical heading as the state
and the heading set point as the input. This model serves as a simplification
of both the turning dynamics of the ornithopter and the on-board PID
controller regulating the actuator inputs to reach the desired orientation. We
use this model in Section~\ref{sec:visual_servoing} to determine the backwards
reachable region for navigating through the window in simulation.
\\
The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. This flight speed is slower than
the maximum speed of the ornithopter to facilitate more reliable tracking.

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
\label{sec:visual_servoing}
%TODO cameron

@@

We determined the backwards reachable set of initial states for successful
window traversal both geometrically and in simulation. The results are shown
in Figure~\ref{fig:feasible_set}. The initial conditions that we chose are
spatially gridded in the camera viewing region at 0.05 meter increments and an
initial angular position perpendicular to the window plane. The geometrically
infeasible region is indicated in blue and was computed using the minimum
turning radius and simple geometry. Both the regions in red and blue indicate
the infeasible region determined in simulation. We simulated the turning model
described in Section~\ref{sec:flight_control}, with a PID controller steering
the system to the center of the window as the input. The infeasible region
calculated in simulation is a superset of the geometrically infeasible region.
The simulation is more realistic as the rotational damping during the turn and
the on-board PID controller cause the real system to react to inputs at a
non-constant angular velocity.
\\
To verify the feasible region experimentally, we conducted trials at initial
distances in X meter increments from the window along the edges of the image
frame of the camera. We started each trial with an initial heading
perpendicular to the window plane. The results of the experiments are shown in
Figure~\ref{fig:flight_paths}.
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
%TODO ryan

@@

%%
%% The following two commands are all you need in the
%% initial runs of your .tex file to
%% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{manuscript}

%% You must have a proper ".bib" file
%%  and remember to run:
%% latex bibtex latex latex
%% to resolve all references
%%
%% ACM needs 'a single self-contained file'!
%%
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%The aamas2013 .cls file is based on the
%sig-alternate.cls file that is itself chock-full of succinct and
%helpful comments.  If you consider yourself a moderately experienced
%to expert user of \LaTeX, you may find reading it useful but please
%remember not to change it.
%%\balancecolumns % GM June 2007
\end{document}

>>>>

@@

% In the original styles from ACM, you would have needed to
% add meta-info here. This is not necessary for AAMAS 2013  as
% the complete copyright information is generated by the cls-files.

\title{Cooperative Control for Narrow Passage Traversal with an 
Ornithopter Micro Air Vehicle and Lightweight Ground Station}

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
%TODO(humphrey)
The flying platform used in this experiment is the H$^2$Bird flapping wing 
MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling 
wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams 
fully loaded. A tail propeller and servo-controlled tail elevator provide 
control in the yaw and pitch axes. An onboard electronics package combines 
a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor 
drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
can operate for approximately [10] minutes.

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
%TODO(ryan)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO(ryan)

@@

%----------------------------------------------------------------------------%
\subsection{Ornithoper Flight Control}
%TODO(humphrey)

\begin{figure}[tb]
\centering
%\includegraphics[width=\linewidth]{figures/h2bird_axes.png}
\includegraphics[height=150pt]{figures/h2bird_axes.png}
\caption{H$^2$Bird Ornithopter with flight control axes}
\label{fig:h2Bird_axes}
\end{figure}

@@

Due to the high pitch angles experienced by the H$^2$Bird, we use a 
quaternion representation of vehicle pose to avoid the numerical 
instabilities of Euler angle representations. This choice of pose 
representation also lends itself well to remote guidance, as relative body 
and world coordinate angular displacements can be represented compactly 
with quaternion multiplication.

Given a reference pose $q$, we define the error $q_e$ as the rotation 
required to reach the reference pose from the measured pose. We then 
linearize the error quaternion to produce body axis angular errors as as 
inputs to the attitude controllers.

@@

\begin{equation}
\label{quat_error}
q_e = q'\otimes q_r
\end{equation}

\begin{equation}
\label{quat_angle}
\alpha = 2acos(q_{e,w})
\end{equation}

\begin{equation}
\label{quat_linearize}
Q_e = \alpha*q_e/sin(\alpha /2)
\end{equation}

%----------------------------------------------------------------------------%
\subsection{Pose Estimation}
%TODO(ryan)

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
%TODO(ryan)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%TODO(ryan)

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.png}
\caption{Conceptual sketch of experimental environment.}
\label{fig:experiment_sketch}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
%TODO(cameron)

To judge the performance of our system described in the previous sections, we
determined the feasible set of initial conditions that resulted in successful
window traversal. We then conducted experiments to calculate the success rate
both within the feasible region and on the edges of the feasible region. The
following sections detail how we calculated the feasible and infeasible
regions using system identification and present the results of our
experimental trials.

%----------------------------------------------------------------------------%
\subsection{Ornithopter Flight Control}
\label{sec:flight_control}
%TODO(cameron)
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
\caption{H$^2$Bird step response}
\label{fig:step_response}
\end{figure}

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. The results of this experiment are depicted in
Figure~\ref{fig:step_response}. The data for the experiment was gathered from
the integration of the H$^2$Bird's built in gyroscope data. We used Matlab to
fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
turning behavior to the step response with the physical heading as the state
and the heading set point as the input. This model serves as a simplification
of both the turning dynamics of the ornithopter and the on-board PID
controller regulating the actuator inputs to reach the desired orientation. We
use this model in Section~\ref{sec:visual_servoing} to determine the backwards
reachable region for navigating through the window in simulation.

The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. This flight speed is slower than
the maximum speed of the ornithopter to facilitate more reliable tracking.

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
\label{sec:visual_servoing}
%TODO(cameron)

@@

We determined the backwards reachable set of initial states for successful
window traversal both geometrically and in simulation. The results are shown
in Figure~\ref{fig:feasible_set}. The initial conditions that we chose are
spatially gridded in the camera viewing region at 0.05 meter increments and an
initial angular position perpendicular to the window plane. The geometrically
infeasible region is indicated in blue and was computed using the minimum
turning radius and simple geometry. Both the regions in red and blue indicate
the infeasible region determined in simulation. We simulated the turning model
described in Section~\ref{sec:flight_control}, with a PID controller steering
the system to the center of the window as the input. The infeasible region
calculated in simulation is a superset of the geometrically infeasible region.
The simulation is more realistic as the rotational damping during the turn and
the on-board PID controller cause the real system to react to inputs at a
non-constant angular velocity.

To verify the feasible region experimentally, we conducted trials at initial
distances in X meter increments from the window along the edges of the image
frame of the camera. We started each trial with an initial heading
perpendicular to the window plane. The results of the experiments are shown in
Figure~\ref{fig:flight_paths}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
%TODO(ryan)

@@

%%
%% The following two commands are all you need in the
%% initial runs of your .tex file to
%% produce the bibliography for the citations in your paper.
%\bibliographystyle{abbrv}
%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this 
% case
%% You must have a proper ".bib" file
%%  and remember to run:
%% latex bibtex latex latex
%% to resolve all references
%%
%% ACM needs 'a single self-contained file'!
%%
%%APPENDICES are optional
%%\balancecolumns
%\appendix
%%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%% This next section command marks the start of
%% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The aamas2013 .cls file is based on the
%sig-alternate.cls file that is itself chock-full of succinct and
%helpful comments.  If you consider yourself a moderately experienced
%to expert user of \LaTeX, you may find reading it useful but please
%remember not to change it.
%%\balancecolumns % GM June 2007
%% That's all folks!
\end{document}
